---
title: "Net Generation"
author: "Dennis Duffy"
date: "2022-09-22"
output: html_document
---

```{r, include = FALSE, echo = FALSE}

library(tidyverse)
library(jsonlite)
library(httr)
library(janitor)
library(downloadthis)
library(data.table)
library(ggrepel)
library(writexl)
library(rio)
library(extrafont)
library(here)
library(keyring)
library(Microsoft365R)

eia_key <- key_get("eia_api")

# SharePoint connection for saving plots
sp_site <- get_sharepoint_site(
  site_url = "https://mn365.sharepoint.com/sites/COMM_EnergyData"
)
sp_drive <- sp_site$get_drive("Documents")

# Helper: save plot locally then upload to SharePoint
save_plot_to_sharepoint <- function(plot, filename, width = 7, height = 5) {
  local_path <- file.path(tempdir(), filename)
  ggsave(local_path, plot, width = width, height = height)
  sp_drive$upload_file(
    src = local_path, 
    dest = paste0("General/EIA Data/Plots/", filename)
  )
  file.remove(local_path)
}

#define date

today <- format(Sys.Date(), format = "%m-%d-%Y")

options(scipen = 999)

raw_generation <- 
  GET(
    paste0(
      "https://api.eia.gov/v2/electricity/electric-power-operational-data/data?api_key=", 
      eia_key, 
      "&data[]=generation&frequency=annual&facets[location][]=MN&frequency=annual&facets[sectorid][]=99"
    )
  )

raw_generation <- fromJSON(rawToChar(raw_generation$content))

gen_data_cleaner <- function(x){
  
  generation_perc <- x$response$data |> 
  tibble() |> 
  arrange(desc(as.numeric(period))) |> 
  clean_names() |> 
  filter(fueltypeid %in% c("BIO", "HYC", "SUN", "WND", "COW", "NG", "PEL", "NUC", "OTH", "ALL", "AOR")) |>
  mutate(fuel_type_description = str_replace_all(fuel_type_description, c("all renewables" = "Renewables", 
                                                                          "natural gas" = "Natural Gas", 
                                                                          "all coal products" = "Coal", 
                                                                          "natural gas" = "Natural Gas",
                                                                          "wind" = "Wind", 
                                                                          "solar" = "Solar", 
                                                                          "nuclear" = "Nuclear", 
                                                                          "all fuels" = "All Fuels", 
                                                                          "biomass" = "Biomass", 
                                                                          "conventional hydroelectric" = "Conventional Hydroelectric", 
                                                                          "other" = "Other", 
                                                                          "petroleum liquids" = "Petroleum Liquids")), 
         generation = as.numeric(generation)) |> 
  select(period, fuel_type_description, generation) |> 
  pivot_wider(names_from = fuel_type_description, values_from = generation) |> 
  mutate(Renewables = Renewables + `Conventional Hydroelectric`) |> 
  mutate(
    across(c(2:12), 
           .fns = ~. / `All Fuels`)
  ) |> 
  select(period, 2:12) |> 
  pivot_longer(!period, names_to = "fuel_type_description", values_to = "generation")

  generation_total <- x$response$data |>
    tibble() |>
    clean_names() |>
    filter(fueltypeid %in% c("BIO", "HYC", "SUN", "WND", "COW", "NG", "PEL", "NUC", "OTH", "ALL", "AOR")) |>
    mutate(fuel_type_description = str_replace_all(fuel_type_description, c("all renewables" = "Renewables",
                                                                            "natural gas" = "Natural Gas",
                                                                            "all coal products" = "Coal",
                                                                            "natural gas" = "Natural Gas",
                                                                            "wind" = "Wind",
                                                                            "solar" = "Solar",
                                                                            "nuclear" = "Nuclear",
                                                                            "all fuels" = "All Fuels",
                                                                            "biomass" = "Biomass",
                                                                            "conventional hydroelectric" = "Conventional Hydroelectric",
                                                                            "other" = "Other",
                                                                            "petroleum liquids" = "Petroleum Liquids")), 
           generation = as.numeric(generation)) |>
    select(period, fuel_type_description, generation) |>
    pivot_wider(names_from = fuel_type_description, values_from = generation) |>
    mutate(Renewables = Renewables + `Conventional Hydroelectric`) |>
    pivot_longer(!period, names_to = "fuel_type_description", values_to = "generation") |>
    left_join(generation_perc, by = c("period", "fuel_type_description")) |>
    rename(total_generation_thousand_mwh = 3,
           percent_generation = 4) |>
    arrange(period, fuel_type_description)
  
  return(generation_total)

}


generation_total <- gen_data_cleaner(raw_generation)

```

View Electricity Generation Data

```{r, echo = FALSE}

data_use <- generation_total |> 
  arrange(desc(period), fuel_type_description)

DT::datatable(data_use, rownames = FALSE)

suppressMessages(
 generation_total |> 
  arrange(period, fuel_type_description) |> 
  download_this(
    output_name = "Electricity Generation in Minnesota", 
    output_extension = ".xlsx", 
    button_label = "Download Data",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```


&nbsp;
&nbsp;

Current Net Generation

This page supplies electricity generation statistics for Minnesota. This shows renewable energy trends, the electricity generation mix in Minnesota. 



Electricity Generation by Source - Most Recent Year
```{r, echo = F}

title_year <- as.character(max(generation_total$period))

data_use <- generation_total |> 
  filter(fuel_type_description %in% c("Renewables", "Coal", "Natural Gas", "Nuclear")) |> 
  mutate(percent_label = paste0(as.character(round(percent_generation * 100)), "%")) |>
  arrange(desc(period))

p <- data_use  |> 
  filter(period == max(period)) |> 
  ggplot() + aes(x = percent_generation, y = fct_reorder(as_factor(fuel_type_description), percent_generation), label = percent_label) + geom_bar(stat = "identity", fill = "#003865") + 
  geom_text(nudge_x = .02) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Percent") +
  ggtitle(paste0("Electricity Generation in Minnesota ", title_year)) + scale_x_continuous(labels = scales::percent, limits = c(0, .4)) + 
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today)) 

  
p

suppressMessages(
  p |> 
  download_this(
    output_name = "Electricity Generation in Minnesota", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

save_plot_to_sharepoint(p, "Electricity Generation in Minnesota.png")
```

&nbsp;
&nbsp;

Net Generation from spreadsheet posted to https://www.eia.gov/electricity/data/state/ (appears to be updated more quickly than the API). Use to verify what gets pulled from API. 

```{r, echo = FALSE}

raw_gen <- rio::import("https://www.eia.gov/electricity/data/state/annual_generation_state.xls", skip = 1) 
  
total_generation <- raw_gen |>
  tibble() |> 
  clean_names() |> 
  filter(state == "MN" & type_of_producer == "Total Electric Power Industry") |> 
  filter(year %in% c((max(raw_gen$YEAR) - 20):max(raw_gen$YEAR))) |> 
  select(year, energy_source, generation_megawatthours) |> 
  pivot_wider(names_from = "energy_source", values_from = "generation_megawatthours") |> 
  rowwise() |> 
  mutate(Renewables = sum(`Wind`, `Wood and Wood Derived Fuels`, `Solar Thermal and Photovoltaic`, `Hydroelectric Conventional`, 
           `Other Biomass`, na.rm = TRUE)) 

total_generation |> 
  mutate(across(c(Coal, `Natural Gas`, Nuclear, Renewables), ~ . / Total)) |> 
  select(year, Coal, `Natural Gas`, Nuclear, Renewables) |> 
  arrange(desc(year))


```


&nbsp;
&nbsp;

Change in Net Generation in Minnesota 
```{r, echo = F}

fuel_colors <- c("Renewables" = "#78BE21", "Nuclear" = "#008EAA", "Coal" = "#000000", "Natural Gas" = "#8D3F2B")

top_year <- max(as.numeric(generation_total$period))

bottom_year <- max(as.numeric(generation_total$period)) - 15

data_use <- generation_total |> 
  mutate(period = as.numeric(period)) |> 
  filter(period == top_year | period == bottom_year) |> 
  select(period, fuel_type_description, percent_generation) |> 
  filter(fuel_type_description %in% c("Renewables", "Coal", "Nuclear", "Natural Gas")) |>
  mutate(percent_label = paste0(as.character(round(percent_generation * 100)), "%"), 
         nudge_factor = ifelse(period == top_year, 1, -1))

p <- data_use |> 
  ggplot() + aes(x = period, y = percent_generation, label = percent_label) + geom_point(aes(colour = factor(fuel_type_description)), size = 3) + geom_line(aes(colour = factor(fuel_type_description)), linewidth = 1) + geom_text_repel(nudge_x = data_use$nudge_factor, size = 3.5) + scale_color_manual(values = fuel_colors, name = "") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11), legend.position = c(.7, .8)) + xlab("Year") +
  ggtitle("Change in Electricity Generation - Minnesota") + scale_y_continuous(labels = scales::percent) + scale_x_continuous(breaks = c(bottom_year, top_year), limits = c((bottom_year - 3), (top_year + 3))) +
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today))

p

suppressMessages(
  p |> 
  download_this(
    output_name = "Change in Electricity Generation - Minnesota", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)


save_plot_to_sharepoint(p, "Change in Electricity Generation - Minnesota.png")

```

&nbsp;
&nbsp;

Change in US Electricity Generation

```{r, echo = FALSE}

raw_generation_us <- 
  GET(
    paste0(
      "https://api.eia.gov/v2/electricity/electric-power-operational-data/data?api_key=", 
      eia_key, 
      "&data[]=generation&frequency=annual&facets[location][]=US&frequency=annual&facets[sectorid][]=99"
      )
  )

raw_generation_us <- fromJSON(rawToChar(raw_generation_us$content))

generation_total_us <-  gen_data_cleaner(raw_generation_us)

top_year <- max(as.numeric(generation_total_us$period))

bottom_year <- max(as.numeric(generation_total$period)) - 15

data_use <- generation_total_us |> 
  mutate(period = as.numeric(period)) |> 
  filter(period == top_year | period == bottom_year) |> 
  select(period, fuel_type_description, percent_generation) |> 
  filter(fuel_type_description %in% c("Renewables", "Coal", "Nuclear", "Natural Gas")) |>
  mutate(percent_label = paste0(as.character(round(percent_generation * 100)), "%"), 
         nudge_factor = ifelse(period == top_year, 1, -1))

p <- data_use |> 
  ggplot() + aes(x = period, y = percent_generation, label = percent_label) + geom_point(aes(colour = factor(fuel_type_description)), size = 3) + geom_line(aes(colour = factor(fuel_type_description)), size = 1) + geom_text_repel(nudge_x = data_use$nudge_factor, size = 3.5) + scale_color_manual(values = fuel_colors, name = "") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Year") +
  ggtitle("Change in Electricity Generation - U.S.") + scale_y_continuous(labels = scales::percent) + scale_x_continuous(breaks = c(bottom_year, top_year), limits = c((bottom_year - 3), (top_year + 3))) +
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today))

p

suppressMessages(
  p |> 
  download_this(
    output_name = "Change in Electricity Generation - U.S.", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)


```



&nbsp;
&nbsp;

Minnesota Electricity Generation by Source (2000-Current)
```{r, echo = F}

bottom_year <- min(as.numeric(generation_total$period))

top_year <- max(as.numeric(generation_total$period))

p <- generation_total |> 
  mutate(period = as.numeric(period)) |> 
  filter(fuel_type_description %in% c("Coal", "Nuclear", "Natural Gas", "Renewables")) |> 
  mutate(percent_label = ifelse(period == top_year, paste0(as.character(round(percent_generation * 100)), "%"), "")) |> 
  ggplot() + aes(x = period, y = percent_generation) + geom_line(aes(colour = factor(fuel_type_description)), size = 1) +
  scale_color_manual(values = fuel_colors, name = "") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Year") + ggtitle("Electricity Generation Trend - Minnesota") + scale_y_continuous(labels = scales::percent) + scale_x_continuous(breaks = seq(bottom_year, top_year, by = 5), limits = c(bottom_year, top_year + 3)) + geom_text_repel(aes(label = percent_label), nudge_x = 1, size = 3.5) + 
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today))

p 

suppressMessages(
  p |> 
  download_this(
    output_name = "Electricity Generation Trend - Minnesota", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

save_plot_to_sharepoint(p, "Electricity Generation Trend - Minnesota.png")

```

&nbsp;
&nbsp;

US Electricity Generation By Source

```{r, echo = F}
raw_us <- 
  GET(
    paste0(
      "https://api.eia.gov/v2/electricity/electric-power-operational-data/data?api_key=", 
      eia_key, 
      "&data[]=generation&frequency=annual&facets[location][]=US&frequency=annual&facets[sectorid][]=99"
      )
  )
 
raw_us <- fromJSON(rawToChar(raw_us$content))

generation_data <- gen_data_cleaner(raw_us)

title_year <- as.character(max(generation_data$period))

data_use <- generation_data |> 
  filter(fuel_type_description %in% c("Renewables", "Coal", "Natural Gas", "Nuclear")) |> 
  mutate(percent_label = paste0(as.character(round(percent_generation * 100)), "%")) |> 
  arrange(desc(period))

p <- data_use  |> 
  filter(period == max(period)) |> 
  ggplot() + aes(x = percent_generation, y = fct_reorder(as_factor(fuel_type_description), percent_generation), label = percent_label) + geom_bar(stat = "identity", fill = "#003865") + 
  geom_text(nudge_x = .02) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Percent") + ggtitle(paste0("Electricity Generation in U.S. ", title_year)) + scale_x_continuous(labels = scales::percent) +
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today))
  
p

suppressMessages(
  p |> 
  download_this(
    output_name = "Electricity Generation in U.S.", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```

&nbsp;
&nbsp;

Midwest Electricity Generation By Source
```{r, echo = F}
raw_wnc <- 
  GET(
    paste0(
      "https://api.eia.gov/v2/electricity/electric-power-operational-data/data?api_key=", 
      eia_key,
      "&data[]=generation&frequency=annual&facets[location][]=WNC&frequency=annual&facets[sectorid][]=99"
      )
  )
 
raw_wnc <- fromJSON(rawToChar(raw_wnc$content))

raw_enc <- 
  GET(
    paste0(
      "https://api.eia.gov/v2/electricity/electric-power-operational-data/data?api_key=", 
      eia_key, 
      "&data[]=generation&frequency=annual&facets[location][]=ENC&frequency=annual&facets[sectorid][]=99"
      )
  )
 
raw_enc <- fromJSON(rawToChar(raw_enc$content))


wnc <- gen_data_cleaner(raw_wnc)

enc <- gen_data_cleaner(raw_enc)

midwest <- suppressMessages(wnc |> 
  bind_rows(enc) |> 
  filter(period == max(period), 
         fuel_type_description %in% c("All Fuels", "Coal", "Renewables", "Natural Gas", "Nuclear")) |> 
  select(period, fuel_type_description, total_generation_thousand_mwh) |> 
  group_by(period, fuel_type_description) |> 
  summarise(total_generation_thousand_kwh = sum(total_generation_thousand_mwh)) |> 
  pivot_wider(names_from = fuel_type_description, values_from = total_generation_thousand_kwh) |> 
  mutate(across(c(Coal, `Natural Gas`, Nuclear, Renewables), ~ . / `All Fuels`)) |> 
  select(-c(`All Fuels`)) |> 
  pivot_longer(!period, names_to = "fuel_type_description", values_to = "percent_generation") |> 
  mutate(percent_label = paste0(as.character(round(percent_generation * 100)), "%")))

title_year = max(midwest$period)


p <- midwest |> 
    ggplot() + aes(x = percent_generation, y = fct_reorder(as_factor(fuel_type_description),   percent_generation), label = percent_label) + geom_bar(stat = "identity", fill = "#003865") + 
    geom_text(nudge_x = .02) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Percent") + ggtitle(paste0("Electricity Generation in Midwest ", title_year)) + scale_x_continuous(labels = scales::percent) + 
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today))

p

suppressMessages(
  p |> 
  download_this(
    output_name = "Electricity Generation in Midwest", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```

MROW Electricity Generation By Source (Hardcoding this for the Quad. See original source here: https://www.epa.gov/egrid/power-profiler#/MROW)

Coal	Gas	Nuclear	Renewables
39.6%	10.6%	8.6%	40.7%

```{r, echo = FALSE}

p <- tribble(
  ~period, ~fuel_type_description, ~percent_generation, ~percent_label, 
  "2021", "Coal", .396, "40%", 
  "2021", "Natural Gas", .106, "11%", 
  "2021", "Nuclear", .086, "9%", 
  "2021", "Renewables", .407, "41%"
) |> 
   ggplot() + aes(x = percent_generation, y = fct_reorder(as_factor(fuel_type_description),   percent_generation), label = percent_label) + geom_bar(stat = "identity", fill = "#003865") + 
    geom_text(nudge_x = .02) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.y = element_blank(), element_text(family = "Calibri", size = 11)) + xlab("Percent") + ggtitle(paste0("Electricity Generation - MROW 2021")) + scale_x_continuous(labels = scales::percent) + 
  labs(caption = "Source: Environmental Protection Agency\nPercentages Exceed 100% Due to Rounding")

p


suppressMessages(
  p |> 
  download_this(
    output_name = "Electricity Generation MROW", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```


&nbsp;
&nbsp;

Total and Renewable Generation

```{r, echo = FALSE}

fuel_colors <- c("Renewables" = "#78BE21", "All Fuels" = "#003865")

top_year <- max(as.numeric(generation_total$period))

bottom_year <- min(as.numeric(generation_total$period))

p <- generation_total |>
  mutate(period = as.numeric(period), 
         gen_label = case_when(
           period == max(period) ~ as.character(round(total_generation_thousand_mwh, 0)), 
           TRUE ~ ""
         )) |> 
  filter(fuel_type_description %in% c("All Fuels", "Renewables"), 
         period >= bottom_year) |> 
  ggplot() + aes(x = period, y = total_generation_thousand_mwh) + geom_line(aes(color = factor(fuel_type_description)), size = 1) + 
  scale_color_manual(values = fuel_colors, name = "") + 
  scale_x_continuous(name = "Year", breaks = seq(bottom_year, top_year, by = 5), limits = c(bottom_year, top_year + 2)) + 
  geom_text_repel(aes(label = gen_label), nudge_x = 1) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), element_text(family = "Calibri", size = 11)) + 
  ggtitle("Minnesota Electricity Generation - Total and Renewables") + 
  labs(caption = paste0("Source: Energy Information Administration\nAccessed ", today)) + 
  ylab("Thousand Megawatthours")
  
p

suppressMessages(
  p |> 
  download_this(
    output_name = "Total and Renewable Generation", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```




