---
title: "R Notebook"
output: html_notebook
---

See report from MISO:https://cdn.misoenergy.org/MISO%20Response%20to%20the%20Reliability%20Imperative504018.pdf

Statewide SAIDI figures retrieved from here: https://www.eia.gov/electricity/annual/html/epa_11_04.html

Statewide SAIFI figures retrieved from here: https://www.eia.gov/electricity/annual/html/epa_11_05.html

```{r, echo = FALSE}

library(tidyverse)
library(rvest)
library(janitor)
library(downloadthis)
library(ggrepel)

url <- "https://www.eia.gov/electricity/annual/html/epa_11_04.html"

reliability_fetch <- function(x){
  
  table_use <- read_html(url) |> 
    html_table() |> 
    pluck(2) |> 
    clean_names() |> 
    select(1:31) |> 
    row_to_names(row_number = 1) |> 
    clean_names()
  
  
  return(table_use)
  
}

med_fetch <- function(x, measure_name){
  
  table_use <- reliability_fetch(x) |> 
    select(1:11) |> 
    row_to_names(row_number = 1) |> 
    rename(region = 1) |> 
    mutate(across(!c(region), as.numeric)) |> 
    pivot_longer(!region, names_to = "year", values_to = measure_name) |> 
    filter(region %in% c("Minnesota", "West North Central", "U.S. Total")) 
  
  return(table_use)
  
}

womed_fetch <- function(x, measure_name){
  
  table_use <- reliability_fetch(url) |> 
    select(1, 12:21) |> 
    row_to_names(row_number = 1) |> 
    rename(region = 1) |> 
    mutate(across(!c(region), as.numeric)) |> 
    pivot_longer(!region, names_to = "year", values_to = measure_name) |> 
    filter(region %in% c("Minnesota", "West North Central", "U.S. Total")) 
  
  return(table_use)
  
}

med <- med_fetch(url, "saidi")

womed <- womed_fetch(url, "saidi")

loss <- reliability_fetch(url) |> 
  select(1, 22:31) |> 
  row_to_names(row_number = 1) |> 
  pivot_longer(!`Census Divisionand State`, names_to = "year", values_to = "saidi_loss_removed") |> 
  rename(region = 1) |> 
  filter(region %in% c("Minnesota", "West North Central", "U.S. Total")) |> 
  mutate(saidi_loss_removed = as.numeric(saidi_loss_removed)) |> 
  left_join(med, by = c("region" = "region", "year" = "year")) |> 
  mutate(supply_disruption = saidi - saidi_loss_removed)

```
&nbsp;
&nbsp;

SAIDI (with Major Event Days) (interpreted as the minutes of non-momentary electric interruptions, per year, the average customer experienced.)

```{r, echo = FALSE}

region_colors <- c("Minnesota" = "#008EAA", "U.S. Total" = "#0D5257", "West North Central Region" = "#8D3F2B")

p <- med |> 
  mutate(region = case_when(
    region == "West North Central" ~ "West North Central Region", 
    TRUE ~ region
  ),
  saidi_label = case_when(
    year == max(year) ~ saidi, 
    TRUE ~ NA
  )) |> 
  ggplot() + aes(x = year, y = saidi, group = region) + geom_line(aes(colour = factor(region)), size = 1) + scale_color_manual(values = region_colors, name = "Region") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab("SAIDI with Major Events") + ggtitle("System Average Interuption Duration Index - With Major Event Days") +
  labs(caption = "Source: Energy Information Administration") + ylim(0, 550) 

p

suppressMessages(
  p |> 
  download_this(
    output_name = "SAIDI With Major Event days", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)


```

&nbsp;
&nbsp;

SAIDI (without Major Event Days) - Interpreted as the number of non-momentary electric interruptions, per year, the average customer experienced.

```{r, echo = FALSE}

p <- womed |> 
  mutate(region = case_when(
    region == "West North Central" ~ "West North Central Region", 
    TRUE ~ region
  )) |> 
  ggplot() + aes(x = year, y = saidi, group = region) + geom_line(aes(colour = factor(region)), size = 1) + scale_color_manual(values = region_colors, name = "Region") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab("SAIDI without Major Events") + ggtitle("System Average Interuption Duration Index - Without Major Event Days") +
  labs(caption = "Source: Energy Information Administration") + ylim(0, 550)

p

suppressMessages(
  p |> 
  download_this(
    output_name = "SAIDI Without Major Event Days", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)
```

```{r, echo = FALSE}

url <- "https://www.eia.gov/electricity/annual/html/epa_11_05.html"

p <- med_fetch(url, "SAIFI") |> 
  mutate(region = case_when(
    region == "West North Central" ~ "West North Central Region", 
    TRUE ~ region
  )) |> 
  rename(Year = year) |> 
  ggplot() + aes(x = Year, y = SAIFI, group = region) + geom_line(aes(colour = factor(region)), size = 1) + scale_color_manual(values = region_colors, name = "Region") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab("SAIFI with Major Events") + ggtitle("System Average Interuption Frequency Index - With Major Event Days") +
  labs(caption = "Source: Energy Information Administration") + ylim(0, 2)

p

suppressMessages(
  p |> 
  download_this(
    output_name = "SAIFI With Major Event Days", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)

```

&nbsp;
&nbsp;

SAIFI Without Major Event Days

```{r, echo = FALSE}

p <- womed_fetch(url, "SAIFI") |> 
  mutate(region = case_when(
    region == "West North Central" ~ "West North Central Region", 
    TRUE ~ region
  )) |> 
  rename(Year = year) |> 
  ggplot() + aes(x = Year, y = SAIFI, group = region) + geom_line(aes(colour = factor(region)), size = 1) + scale_color_manual(values = region_colors, name = "Region") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab("SAIFI without Major Events") + ggtitle("System Average Interuption Frequency Index - Without Major Event Days") +
  labs(caption = "Source: Energy Information Administration") + ylim(0, 2)

p

suppressMessages(
  p |> 
  download_this(
    output_name = "SAIFI Without Major Event Days", 
    output_extension = ".png", 
    button_label = "Download Plot",
    button_type = "success", 
    has_icon = TRUE,
    icon = "fa fa-save"
  )
)


```


